---
title: "ipums_data5"
format:
  html:
    theme: default
---

Script to compare different models in the discriminant-based metrics for synthesized data with and without permuting gender
```{r}
library(tidyverse)
library(ipumsr)
library(srvyr)
library(tidysynthesis)
library(parsnip)
library("data.table")
library(recipes)
library(devtools)
load_all('../syntheval')
library(syntheval)
library(dials)
library(tune)
library(gtools)
```

```{r}
data <- read_ipums_micro("data/usa_00004.xml")
data <- data |>
  mutate(across(where(is.labelled), ~as_factor(lbl_clean(.x))))
```


```{r}
states <- c("Michigan")
vars <- c("STATEFIP", "PERNUM", "SEX", "AGE", "RACE", 
           "HCOVANY", "EDUC", "EMPSTAT", "FTOTINC", "POVERTY",
           "VETSTAT")

```

```{r}
data_mi <- data %>%
  select(all_of(vars)) %>%
  filter(STATEFIP %in% states,
         PERNUM == 1) 
# recode factor variables (vetstat, sex, race, healthcare coverage, employment status didn't need recoding)
data_mi$EDUC <- recode(data_mi$EDUC, 
               "N/A or no schooling" = "N/A",
               "Nursery school to grade 4" = "Less than high school",
               "Grade 5, 6, 7, or 8" = "Less than high school",
               "Grade 9" = "Some high school",
               "Grade 10" = "Some high school",
               "Grade 11" = "Some high school",
               "Grade 12" = "High school",
               "1 year of college" = "Some college",
               "2 years of college" = "Some college",
               "4 years of college" = "4+ years of college",
               "5+ years of college" = "4+ years of college")

# recode numeric variables (age, poverty (as % of poverty threshold), and income)
data_mi$AGE <- as.numeric(as.character(recode(data_mi$AGE, 
                      "Less than 1 year old"= "0",
                      "90 (90+ in 1980 and 1990)" = "90")))
data_mi$FTOTINC <- as.numeric(as.character(recode(data_mi$FTOTINC,
                                     "No income (1950-2000, ACS/PRCS) " = "0")))
data_mi$POVERTY <- as.numeric(as.character(recode(data_mi$POVERTY,
                                     "501 percent or more of poverty threshold" = "500",
                                     "1 percent or less of poverty threshold (including 0 or negative income)" = "0")))

# set character NAs to be NA
data_mi[data_mi == "N/A"]<-NA 

# complete case analysis of adults only
data_mi <- data_mi %>%
  filter(AGE >= 19,
         complete.cases(.)) %>%
  select(-c("STATEFIP", "PERNUM"))

# get rid of unused NA levels in factors (educ, empstat, vetstat)
data_mi = droplevels(data_mi)
```

```{r}
table(data_mi$SEX, data_mi$RACE)
prop.table(table(data_mi$SEX, data_mi$HCOVANY), 1)
prop.table(table(data_mi$SEX, data_mi$EMPSTAT), 1)
prop.table(table(data_mi$SEX, data_mi$EDUC), 1)
prop.table(table(data_mi$SEX, data_mi$VETSTAT), 1)

ggplot(data_mi, aes(group = SEX, x = log(FTOTINC), color = SEX))+
  geom_density()

ggplot(data_mi, aes(group = SEX, x = POVERTY, color = SEX))+
  geom_density()

data_mi <- data_mi %>%
  select(-c(RACE))
```

Synthesize data from permuted and original confidential data
```{r}
# original synthesis
start_data = data_mi[,1]
conf_data = data_mi
# synthesize categorical first, then continuous
visit_sequence = visit_sequence(c("VETSTAT", "EMPSTAT", "EDUC", "HCOVANY", "POVERTY", "FTOTINC", "AGE"), start_data, type = "manual")
roadmap = roadmap(conf_data, start_data, visit_sequence)
recipe = construct_recipes(roadmap = roadmap)

tree_cl <- parsnip::decision_tree(cost_complexity = .0001) %>%
  set_mode(mode = "classification") %>%
  set_engine(engine = "rpart")
tree_reg <- parsnip::decision_tree(cost_complexity = .0001) %>%
  set_mode(mode = "regression") %>%
  set_engine(engine = "rpart")

synth_spec = synth_spec(roadmap,
                        synth_algorithms = list(tree_cl, tree_cl, tree_cl, tree_cl, tree_reg, tree_reg, tree_reg),
                        recipe,
                        predict_methods = sample_rpart)

# noise
noise <- noise(roadmap = roadmap,
               add_noise = FALSE,
               exclusions = 0)

# constraints
constraints <- constraints(roadmap = roadmap,
                           constraints = NULL,
                           max_z = 0)

replicates <- replicates(replicates = 1,
                         workers = 1,
                         summary_function = NULL)

# create a presynth object
presynth1 <- presynth(
  roadmap = roadmap,
  synth_spec = synth_spec,
  noise = noise, 
  constraints = constraints,
  replicates = replicates
)
```

```{r}
set.seed(1)
synthesized = synthesize(presynth1, progress = TRUE)
synthesized_perm = synthesize(presynth1, progress = TRUE)
synthesized_perm$synthetic_data$SEX = permute(data_mi$SEX)
```

```{r}
disc_tree <- function(discriminator){
  # Evaluate discriminant-based metrics on the data using tree-based model
  tree_mod <- decision_tree(cost_complexity = tune()) %>%
    set_mode(mode = "classification") %>%
    set_engine(engine = "rpart")
  
  rpart_rec <- recipe(.source_label ~ ., data = discriminator$combined_data)
  
  grid = grid_regular(cost_complexity(), levels= 10)
  
  # set up discriminator
  d <- discriminator %>%
    add_propensities_tuned(
      grid = grid, 
      recipe = rpart_rec,
      spec = tree_mod
    ) %>%
    add_discriminator_auc() %>%
    add_specks() %>%
    add_pmse()
  res = c(d$discriminator_auc, d$pmse, d$specks)
  return(c(res$.estimate[2], res$.pmse[2], res$.specks[2]))
}

disc_rf <- function(discriminator){
  # Evaluate discriminant-based metrics on the data using random forest
  rand_forest_mod <- rand_forest(
    mode = "classification",
    engine = "randomForest",
    mtry = NULL,
    trees = NULL,
    min_n = NULL
  )
  
  rpart_rec <- recipe(.source_label ~ ., data = discriminator$combined_data)
  
  # set up discriminator
  d_rf <- discriminator %>%
    add_propensities(
      recipe = rpart_rec,
      spec = rand_forest_mod
    ) %>%
    add_discriminator_auc() %>%
    add_specks() %>%
    add_pmse()
  res = c(d_rf$discriminator_auc, d_rf$pmse, d_rf$specks)
  return(c(res$.estimate[2], res$.pmse[2], res$.specks[2]))
}

disc_lasso <- function(discriminator){
  # Evaluate discriminant-based metrics on the data using random forest
  lasso_mod <- logistic_reg(penalty = tune(), mixture = 1) %>%
      set_mode(mode = "classification") %>%
      set_engine(engine = "glmnet")  
  
  lasso_rec <- recipe(.source_label ~ ., data = discriminator$combined_data) %>%
    step_dummy(all_nominal_predictors()) 
  lasso_rec <- lasso_rec %>%
    step_interact(terms = ~ SEX_Female:(starts_with("EDUC")+
                                          starts_with("EMPSTAT")+
                                          all_numeric_predictors()), keep_original_cols = TRUE)
  
  # make sure interaction worked correctly
  colnames(bake(prep(lasso_rec, training = discriminator$combined_data), 
                discriminator$combined_data))
  
  # set up discriminator
  grid = grid_regular(penalty(), levels = 20)
  # set up discriminator
  d_lasso <- discriminator %>%
    add_propensities_tuned(
      grid = grid,
      recipe = lasso_rec,
      spec = lasso_mod
    ) %>%
    add_discriminator_auc() %>%
    add_specks() %>%
    add_pmse()
  
  res = c(d_lasso$discriminator_auc, d_lasso$pmse, d_lasso$specks)
  return(c(res$.estimate[2], res$.pmse[2], res$.specks[2]))
  #return(d_lasso)
  #library(vip)
  
  #d_lasso$discriminator %>%
  #  extract_fit_parsnip() %>%
  #  vip()
}
```


```{r}
# data synthesized from confidential data
d_synthesized = discrimination(synthesized$synthetic_data, data_mi)
d1 = disc_tree(d_synthesized)
d2 = disc_rf(d_synthesized)
d3 = disc_lasso(d_synthesized)

# data synthesized from permuted confidential data
d_synthesized_perm = discrimination(synthesized_perm$synthetic_data, data_mi)
d4 = disc_tree(d_synthesized_perm)
d5 = disc_rf(d_synthesized_perm)
d6 = disc_lasso(d_synthesized_perm)

c(d1, d2, d3, d4, d5, d6)
```

Running these synthesis procedures on the women and man subsets of the data (starting with just the women subset)
```{r}
synthesized_f <- synthesized$synthetic_data %>%
  filter(SEX == "Female")
data_mi_f <- data_mi %>%
  filter(SEX == "Female")

d_synthesized_f = discrimination(synthesized_f, data_mi_f)
d7 = disc_tree(d_synthesized_f)
d8 = disc_rf(d_synthesized_f)
d9 = disc_lasso(d_synthesized_f)

synthesized_perm_f <- synthesized_perm$synthetic_data %>%
  filter(SEX == "Female")
d_synthesized_perm_f <- discrimination(synthesized_perm_f, data_mi_f)
d10 = disc_tree(d_synthesized_perm_f)
d11 = disc_rf(d_synthesized_perm_f)
d12 = disc_lasso(d_synthesized_perm_f)

c(d7, d8, d9, d10, d11, d12)
```

Investigating pairwise utility -- want to see whether trying to get utility of sex with other variables (using a decision tree, which is generally best, and AUC), pulls out that utility of sex with other thigns is worse in the permuted dataset than the non-permuted one.
```{r}
disc_tree_pair <- function(discriminator, pair_var){
  aucs = c()
  # Evaluate discriminant-based metrics on the data using tree-based model
  tree_mod <- decision_tree(cost_complexity = tune()) %>%
    set_mode(mode = "classification") %>%
    set_engine(engine = "rpart")
  
  rpart_rec <- recipe(.source_label ~ ., data = discriminator$combined_data)
  
  grid = grid_regular(cost_complexity(), levels= 10)
  
  for (var in colnames(discriminator$combined_data)[-1]){ # iterate through everything but source.label
    d_temp = discriminator
    d_temp$combined_data = d_temp$combined_data %>%
      select(all_of(c(".source_label", var, pair_var)))
    # set up discriminator
    d <- discriminator %>%
      add_propensities_tuned(
        grid = grid, 
        recipe = rpart_rec,
        spec = tree_mod
      ) %>%
      add_discriminator_auc()
    aucs = append(aucs, d$discriminator_auc$.estimate[2])
  }
  return(aucs)
}
```

```{r}
d_synthesized = discrimination(synthesized$synthetic_data, data_mi)
aucs = disc_tree_pair(d_synthesized, "SEX")

d_synthesized_perm = discrimination(synthesized_perm$synthetic_data, data_mi)
aucs_perm = disc_tree_pair(d_synthesized_perm, "SEX")

aucs
aucs_perm
```
